---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tceqcrp

<!-- badges: start -->
[![R-CMD-check](https://github.com/mps9506/tceqcrp/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/mps9506/tceqcrp/actions/workflows/R-CMD-check.yaml)
[![R-universe version](https://mps9506.r-universe.dev/tceqcrp/badges/version)](https://mps9506.r-universe.dev/tceqcrp)
<!-- badges: end -->


Scripted downloads from the TCEQ Clean Rivers Program (CRP) water
quality data portal: https://www80.tceq.texas.gov/SwqmisWeb/public/crpweb.faces#.

## Installation

tceqcrp is not currently on CRAN, but can be installed from r-universe:

```r
install.packages('tceqcrp', repos = c('https://mps9506.r-universe.dev', 'https://cloud.r-project.org'))
```


## Quick Start

### Concepts

- **Basin**: a major river basin (1-25), e.g. Brazos River.
- **Segment**: a subdivision within a basin (e.g. `1202`, "Brazos River
  Below Navasota River"). You select which segments to download.
- **Station**: an individual monitoring site (e.g. `11846`). Stations
  appear in the `Station ID` / `Station Description` columns of the
  downloaded data — they are not selected directly.

```{r}
library(tceqcrp)

# See what's available
crp_basins()
crp_data_types()

# List segments in the Brazos River basin
segments <- crp_list_segments(basin = 12)
head(segments)

# Download Field Water Quality data for two stations
brazos <- crp_query(
  basin = 12,
  startdate = "01/01/2020",
  enddate = "12/31/2020",
  data_type = 1, # Field Water Quality
  segment_ids = c("1201", "1202"),
  clean_names = TRUE
)
brazos
```

## How it works

The CRP portal is a JavaServer Faces (JSF) application. This package:

- Opens a session and captures the JSF ViewState token.
- Selects a river basin (which repopulates the station list server-side).
- Submits the query and follows the redirect to StreamServlet,
    which streams a tab-delimited data file.

Session cookies are preserved across all requests so the server keeps
the query state consistent.


## Disclaimer

This package accesses a public TCEQ website by mimicking browser form
submissions. It is not affiliated with or endorsed by TCEQ. Please use
responsibly and review TCEQ's terms of use.
